---
title: "Week 10 Lab: Applications of Tidyverse Skills"
from: markdown+emoji
subtitle: AAI1001 -- Data Engineering and Visualization<br>AY24/25 Tri 2 (Assisted by GenAI)<br>Exercise 5 (Teamwork 2)
author:  
  name: Team [Insert your Team ID]
  affiliations: Singapore Institute of Technology
date: 2025-03-11
date-format: "dddd MMM D, YYYY"
format: 
  html: 
    toc: true
    toc-depth: 1
    toc-location: left
    number-sections: true
    number-depth: 3
    embed-resources: true
    html-table-processing: none
---

```{r}
#| label: setup
#| include: false

library(tidyverse)
library(gt)
```


# Converting Football Scores into a League Table (30%)


## Generating a League Table
```{r}
epl_results <- read_csv("epl_results.csv", show_col_types=FALSE) |>
  as_tibble()

epl_teams <- read_csv("epl_teams.csv", show_col_types=FALSE) |>
  as_tibble()
```

```{r}
library(tidyverse)

# Load Data
epl_results <- read_csv("epl_results.csv", show_col_types = FALSE)
epl_teams <- read_csv("epl_teams.csv", show_col_types = FALSE)

# Convert match results to separate columns for home and away goals
epl_results_long <- epl_results %>%
  pivot_longer(-home_team, names_to = "away_team", values_to = "score") %>%
  separate(score, into = c("home_goals", "away_goals"), sep = "-", convert = TRUE)

# Compute match points and statistics
match_results <- epl_results_long %>%
  mutate(
    home_points = case_when(
      home_goals > away_goals ~ 3,
      home_goals == away_goals ~ 1,
      TRUE ~ 0
    ),
    away_points = case_when(
      home_goals < away_goals ~ 3,
      home_goals == away_goals ~ 1,
      TRUE ~ 0
    )
  )

# Compute cumulative stats for each team
home_stats <- match_results %>%
  group_by(home_team) %>%
  summarise(
    points = sum(home_points),
    goals_scored = sum(home_goals)
  )

away_stats <- match_results %>%
  group_by(away_team) %>%
  summarise(
    points = sum(away_points),
    goals_scored = sum(away_goals)
  )

# Merge home and away stats
total_stats <- full_join(
  home_stats %>% rename(team = home_team),
  away_stats %>% rename(team = away_team),
  by = "team", suffix = c("_home", "_away")
) %>%
  mutate(
    points = points_home + points_away,
    goals_scored = goals_scored_home + goals_scored_away
  ) %>%
  select(team, points, goals_scored)

# Compute goal difference
total_stats <- match_results %>%
  group_by(home_team) %>%
  summarise(goals_conceded_home = sum(away_goals)) %>%
  full_join(
    match_results %>%
      group_by(away_team) %>%
      summarise(goals_conceded_away = sum(home_goals)),
    by = c("home_team" = "away_team")
  ) %>%
  rename(team = home_team) %>%
  mutate(
    goals_conceded = coalesce(goals_conceded_home, 0) + coalesce(goals_conceded_away, 0)
  ) %>%
  right_join(total_stats, by = "team") %>%
  mutate(goal_diff = goals_scored - goals_conceded) %>%
  select(team, points, goal_diff, goals_scored)

# Rank teams based on points, goal difference, and goals scored
epl_league_table <- total_stats %>%
  arrange(desc(points), desc(goal_diff), desc(goals_scored)) %>%
  mutate(rank = row_number())

# Print the league table
epl_league_table

```

```{r}
library(dplyr)
library(tidyr)
library(readr)
library(purrr)

# Read CSV files
epl_results <- read_csv("epl_results.csv", show_col_types = FALSE)
epl_teams <- read_csv("epl_teams.csv", show_col_types = FALSE)

# Function to extract match details safely
calculate_points <- function(home_goals, away_goals) {
  if (is.na(home_goals) || is.na(away_goals)) return(NA)  # Handle missing values
  if (home_goals > away_goals) return(3)  # Home win
  if (home_goals < away_goals) return(0)  # Home loss
  return(1)  # Draw
}

# Convert wide format to long format, removing missing values
results_long <- epl_results %>%
  pivot_longer(cols = -home_team, names_to = "away_team", values_to = "score") %>%
  drop_na(score) %>%  # Remove rows where score is missing
  separate(score, into = c("home_goals", "away_goals"), sep = "-", convert = TRUE)

# Home team statistics
home_stats <- results_long %>%
  mutate(
    points = map2_int(home_goals, away_goals, ~ ifelse(is.na(.x) | is.na(.y), 0, calculate_points(.x, .y))),
    goal_diff = home_goals - away_goals,
    goals_scored = home_goals
  ) %>%
  group_by(home_team) %>%
  summarise(
    points = sum(points, na.rm = TRUE),
    goal_diff = sum(goal_diff, na.rm = TRUE),
    goals_scored = sum(goals_scored, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  rename(team = home_team)

# Away team statistics
away_stats <- results_long %>%
  mutate(
    points = map2_int(away_goals, home_goals, ~ ifelse(is.na(.x) | is.na(.y), 0, calculate_points(.y, .x))),  # Reverse order
    goal_diff = away_goals - home_goals,
    goals_scored = away_goals
  ) %>%
  group_by(away_team) %>%
  summarise(
    points = sum(points, na.rm = TRUE),
    goal_diff = sum(goal_diff, na.rm = TRUE),
    goals_scored = sum(goals_scored, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  rename(team = away_team)

# Combine home and away stats
league_table <- bind_rows(home_stats, away_stats) %>%
  group_by(team) %>%
  summarise(
    points = sum(points, na.rm = TRUE),
    goal_diff = sum(goal_diff, na.rm = TRUE),
    goals_scored = sum(goals_scored, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  left_join(epl_teams, by = c("team" = "abbr")) %>%
  select(team = name, points, goal_diff, goals_scored) %>%
  arrange(desc(points), desc(goal_diff), desc(goals_scored)) %>%
  mutate(rank = row_number()) %>%
  select(rank, team, points, goal_diff, goals_scored)

# Print the final league table
league_table
```




## Improving the Visualization Using gt Functions



# Finding and Visualizing a Real-World Data Set (30%)


## Explanation of the Data Set


## Validation of Conditions C, D, and E


## Generating a Scatter Plot Using `geom_point()`


## Producing and Discussing Two Distinct Plots Designed to Mitigate Overplotting


## Creating Two Distinct Smoothing Interpolations



# Visualizing Michelson’s Speed-of-Light Measurements (25%)


## Converting the `Expt` Column to a Factor


## Scatter Plot: Experiment Number versus Measured Speeds


## Creating a Box Plot with One Box per Experiment


## Inserting a Violin Plot into the Box Plot


## Creating a Faceted Plot Displaying Histograms


## Discussion of the Plots



# Reflection on the Use of GenAI (10%)