---
title: "Week 10 Lab: Applications of Tidyverse Skills"
from: markdown+emoji
subtitle: AAI1001 -- Data Engineering and Visualization<br>AY24/25 Tri 2 (Assisted by GenAI)<br>Exercise 5 (Teamwork 2)
author:  
  name: Team [Insert your Team ID]
  affiliations: Singapore Institute of Technology
date: 2025-03-11
date-format: "dddd MMM D, YYYY"
format: 
  html: 
    toc: true
    toc-depth: 1
    toc-location: left
    number-sections: true
    number-depth: 3
    embed-resources: true
    html-table-processing: none
---

```{r}
#| label: setup
#| include: false

library(tidyverse)
library(gt)
```


# Converting Football Scores into a League Table (30%)


## Generating a League Table
```{r}
epl_results <- read_csv("epl_results.csv", show_col_types=FALSE) |>
  as_tibble()

epl_teams <- read_csv("epl_teams.csv", show_col_types=FALSE) |>
  as_tibble()
```


## Improving the Visualization Using gt Functions




# Finding and Visualizing a Real-World Data Set (30%)


## Explanation of the Data Set


## Validation of Conditions C, D, and E


## Generating a Scatter Plot Using `geom_point()`
```{r}
# library(ggplot2)
# library(dplyr)
# 
# resale_flats <- read.csv("ResaleFlatPrices/Resale Flat Prices (Based on Approval Date), 1990 - 1999.csv")
# 
# # Display the structure of the dataset
# str(resale_flats)
# 
# # Check the number of observations
# num_observations <- nrow(resale_flats)
# cat("Number of observations:", num_observations, "\n")
# 
# # Scatter plot using geom_point()
# ggplot(resale_flats, aes(x = floor_area, y = resale_price)) +
#   geom_point() +
#   labs(
#     title = "Scatter Plot of Floor Area vs. Resale Price",
#     x = "Floor Area (sqm)",
#     y = "Resale Price (SGD)",
#     caption = "Data Source: Singapore Department of Statistics"
#   ) +
#   theme_minimal()

```



## Producing and Discussing Two Distinct Plots Designed to Mitigate Overplotting


## Creating Two Distinct Smoothing Interpolations



# Visualizing Michelson’s Speed-of-Light Measurements (25%)


## Converting the `Expt` Column to a Factor


## Scatter Plot: Experiment Number versus Measured Speeds


## Creating a Box Plot with One Box per Experiment


## Inserting a Violin Plot into the Box Plot


## Creating a Faceted Plot Displaying Histograms


## Discussion of the Plots



# Reflection on the Use of GenAI (10%)